Generalized measurement invariance tests with application to factor analysis

### 2011-11-08 www/paper.Rnw
Addressed a number of reviewer comments pertaining to factor mixture models, moderated factor models, and some other miscellaneous issues.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@71 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Use in Practice}
% Simultaneous test of Meredith's various types of invariance
The proposed tests give researchers a new set of tools for studying 
measurement invariance.  For example, they give researchers the
flexibility to: (1)~simultaneously test all model parameters, yielding
results relevant to many types of measurement
invariance (see, e.g., \citeNP{Mer93}), or (2)~test a single subset of
model parameters, potentially leading to improved power to detect a
single type of measurement invariance.  The traditional
steps have involved hypothesizing a specific type of invariance and
then testing for it via a LR~test, but this is
unnecessary under the proposed framework.

% Researchers interpreting subgroups violating MI
In addition to simultaneously testing different types of measurement
invariance, the proposed tests allow researchers to easily interpret
the nature of the invariance violation.  This is made possible through
the tests' abilities to estimate $\nu$, the threshold dividing
individuals into subgroups that violate
measurement invariance.  While a single $\nu$ was assumed in this
paper, it is also possible to define 
formal rules for estimating multiple $\nu$ parameters
(see \citeNP{ZeiSha10}).  We plan to explore these issues in the future.

>>>>

\subsection{Use in Practice}
The proposed tests give researchers a new set of tools for studying 
measurement invariance, allowing them the 
flexibility to: (1)~simultaneously test all model parameters across
all individuals, yielding
results relevant to many types of measurement
invariance (see, e.g., \citeNP{Mer93}), (2)~test a subset of
model parameters, either across all individuals or within a
multiple-group model, and (3)~use the tests as a type of modification
index following significant LR~tests.
Traditional steps to studying measurement invariance have involved
sequential LR~tests for various types of invariance using
multiple-group models.  As shown in the Example section, it can be
beneficial to employ the traditional steps in tandem with the proposed
tests.  Further, when groups are not defined in advance (e.g.,
continuous $V$), the traditional steps fail unless further assumptions
about the nature of the groups are made.

In addition to providing general information about whether or not
measurement 
invariance holds, the proposed tests allow researchers to interpret
the nature of the invariance violation.  This is made possible, e.g.,
through the tests' abilities to estimate $\nu$, the threshold dividing
individuals into subgroups that violate
measurement invariance (see Equation~\eqref{eq:h1*}).  It is also
possible to define formal rules for estimating multiple $\nu$
parameters (see \citeNP{ZeiSha10}).

\subsection{Comparison to Other Methods}
The step
functions from~\eqref{eq:h18} were employed in the current paper to
highlight connections with nested model
comparison, but the proposed tests typically have non-trivial power for
all (non-constant) deviation patterns $\theta(v_i)$ (considered in
Equation~\eqref{eq:h1}). Thus, the tests will also have power for linear
deviations from parameter constancy, although other techniques may 
perform better in this particular case.  Examples include the 
class of moderated factor models \cite{BauHus09,MolDol10,Pur02}, 
whereby continuous moderators are allowed to linearly impact model
parameters (Purcell also considers quadratic effects of the
moderators).   
Returning to deviations in parameter constancy, if the change in
parameters is continuous but not exactly linear - e.g., a sigmoidal
shift from one set of parameters to another one - then the simple
single shift model in~\eqref{eq:h1*} may be a useful first
approximation and have better power than linear
techniques.

% paragraph on factor mixture models
The proposed tests are also similar to factor mixture models (e.g.,
\citeNP{DolMaa98,LubMut05}) in that they can handle unknown subgroups
violating measurement invariance.  Covariates can also be used to
predict the unknown subgroups under both approaches.
For example, for the 
data used in our simulations, one could employ a factor mixture
model with two latent classes.  This would typically involve the
specification that student age has a linear effect on logit(P(Class
2)), which highlights the main difference between the two techniques:
the factor mixture model requires a specific assumption about the type
of deviation from parameter constancy, while the proposed tests do
not.

More generally, all the approaches described above can be distinguished
from the proposed tests in that they require estimation of a new model
with increased complexity (due to the moderator/covariate).
The proposed tests, on the other hand, are of the
``posthoc'' variety, relying only on results calculated during the
original model estimation.  While no method clearly dominates 
across all situations, use of the proposed tests can at least reduce 
some of the technical issues associated with estimating and
interpreting models of greater complexity.  This may not be a
good reason to use the tests, but it is a consideration that is often
meaningful in practice.
